Spatial distribution of naming patterns in Russian law firms through text embeddings

### 2025-02-28 law-firms-names/journal-article/Paper.Rmd
Minor adjustments before publication, #4

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  warning = FALSE, 
  message = FALSE, 
  echo = FALSE, 
  fig.width = 6.7, 
  fig.id = FALSE,
  dpi = 1200)
IS_PAPER <- TRUE # used in script to disable saving of names and samples
```

```{r package-imports, include=FALSE}
library(dplyr)
library(ggplot2)
library(glue)
library(here)
library(knitr)
library(readr)
library(readxl)
library(sf)
library(stringr)
library(tidyr)
library(units)
```

@@

Russia belongs to the Germanistic branch of a civil law legal system. Legal services (advice, negotiation, representation in court, etc) for organisations and individuals are usually provided by private law companies and individual lawyers. The market is divided between ordinary lawyers and advocates (members of the Bar), but there is no bar monopoly. There are ~90, 000 law firms in Russia, including ~40, 000 commercial law companies, ~30, 000 sole practitioners, ~20, 000 bar (advocate) offices, and ~10, 000 bar (advocate) organisations. There are no geographical restrictions on legal practitioners, so that a law firm registered in one region of the country can provide services in any other region. There are full-service and specialised law firms, as well as Russian law firms (the majority) and foreign law firms (the influential minority).

@@

The issue of obtaining information about firm names is a challenging one. All Russian organizations are registered in the state registry operated by the Federal Tax Service (FTS of Russia). An extract from this registry about a particular company can be searched and downloaded online by any person free of charge. However, a considerable sum of money must be paid and a written request submitted to the FTS of Russia to purchase a complete copy of the registry. Commercial companies offer data resell services, but the prices are similarly high.

To address these challenges, a unified state registry of small and medium-sized enterprises (SME registry) may be used [@SMERegistry2024]. The registry has been maintained since 2016 in accordance with Article 4.1 of the Federal Law of 24 July 2007 No. 209-FZ “On the development of small and medium-sized businesses in the Russian Federation”. As a rule of thumb, all companies and individual entrepreneurs with a yearly revenue below 2 billion roubles and an average yearly number of employees below 250 are automatically included in the SME registry, without any action being required on their part. There are additional qualifications for the SME registry, such as the absence of non-SME or foreign control, but these are unlikely to affect a significant proportion of law firms. The registry is fully downloadable at the FTS open data portal at no cost. Additionally, the portal provides a monthly record of updates made since the registry's inception. To sum it up, the SMB registry can be regarded as the most comprehensive free data source offering information on Russian business entities, both organisations and individuals.

@@

For a validation, a list of law firms' unique tax IDs obtained from the SME registry for the year 2023 is matched with the list of law firms' tax IDs from the Russian Financial Statements Database (RFSD) [@Bondarkov2025] for the same year. The comparison shows that ~3,300 of the firms present in the SME registry are missing from the RFSD, and ~5,200 firms present in the RFSD are missing from the SME registry. The balance of missing records is not so far from being equal. The SME registry contains ~34000 law firms for the year 2023, so the missing rate is below 15%. Consequently, the completeness of the data is relatively high and reliable conclusions can be drawn.

@@

In this study, the term “law firm” is defined as a commercial organization registered in Russia in accordance with the Civil Code of Russia, with a NACE Rev. 2-compatible main activity code equal to 69.10 (activity in the area of law). The aforementioned definition excludes individual entrepreneurs, foreign companies, non-profit organizations, bar (advocacy) formations, and notaries. The selection technique based on the main activity code is derived from that proposed by @Moiseeva2016. Individual entrepreneurs are excluded from the analysis because they typically operate under their personal names, and even if they have a brand, it is not listed in the SME registry. Other types of law firms are not included in the analysis primarily due to data availability issues.

@@

The Yandex Foundation Models cloud service [@Yandex2024] is used to transform the firm names into text vectors, or embeddings. The “text-search-query” model is selected because its purpose, as stated in the documentation, is to vectorize short texts. Prior to vectorization, the firm names are preprocessed to remove indications of organizational form such as “limited liability company”. This is done because such indications are common to the majority of names and thus have little value. The results of the queries to the Yandex API are stored separately for reuse, given that API calls are paid and take time to be completed. The dimension of a firm name vector is 256. Contextual embeddings (YandexGPT) were used in preference to static ones (e.g. FastText), because in the preliminary analysis, they have resulted in a much better clustering of law firm names, which is crucial for naming patterns identification.

@@

The resulting clusters are then manually grouped. A random sample of 20 names is selected from each cluster. With 50 clusters and 20 sampled names allocated in each, the total size of the dataset for manual processing is 1, 000. The cluster samples are carefully examined, and a concise description is provided for each, e.g. “Names with the word “консалт/консалтинг” (consult/consulting)”. After this, the clusters are examined again and assigned to a naming pattern, e.g. “Consulting” or “Law firm”. Finally, the clusters and their naming patterns are divided into four previously mentioned naming strategies derived from Bassols i Gardella et al. [@BassolsiGardella2023]. Each cluster belongs to a single pattern, and each pattern belongs to a single strategy. Each strategy consists of several patterns, and each pattern consists of one or many clusters, forming a hierarchical structure.

@@

Spatial autocorrelation analysis is used to check whether the spatial distribution of names following each pattern identified in the previous step is random or has some regularities. Regions (constituent entities) of the Russian Federation are chosen as the unit of observation. Municipal territories, the next level of spatial division, are not used, because an exploratory data analysis has shown that the lion's share of such the spatial units do not have any law firms.

Six spatial weight matrices are utilized to ensure the robustness of the results. Three of these matrices are constructed using the K-nearest neighbors technique, where K is 3, 4 or 6. Another matrix relies on the existence of common boundaries or points to detect neighboring regions. The Sakhalin and Kaliningrad regions, which are enclaves, are manually linked to a few nearest regions. Two remaining matrices treat regions belonging to the same upper-level macro-region as neighbors, and all other regions as non-neighbors. One of these matrices uses economic regions act as macro-regions, and the other uses federal districts. Economic regions are defined according to the classification that has been widely used by economic geographers and spatial development practitioners in Russia for decades, as outlined, for example, in [@Leyzerovich2010]. The composition of federal districts is defined by the Decree of the President of the Russian Federation (No. 849 of May 13, 2000) “On the Plenipotentiary Representative of the President of the Russian Federation in the Federal District”, with respective amendments.

@@

## 3.1. Law firms naming patterns and strategies

After clustering and manual grouping, 22 naming patterns are identified and assigned to one of four naming strategies (see Fig. 1 for a dendrogram and Table 1 for a sample of 10 clusters, their descriptions, patterns, and strategies). The most popular is the no-strategy followed by ~18, 700 law firms. Slightly less important is the belonging strategy with ~15, 000 companies. The uniqueness strategy (~8, 300 law firms) is relatively rare, and the least used is the market strategy (~5, 300 companies).

@@

```{r table-1-clusters}
clusters_info_table %>% 
  slice_head(n = 10) %>% 
  kable(
    col.names = c(
      "# of cluster", "# of names", "Description",
      "Pattern", "Strategy"
    ),
    caption = "**Table 1.** Characteristics of clusters of law firm names"
  )
write_csv(clusters_info_table, "ESM2.csv")
```

Market strategy, according to the original definition, is used by firms “having a name directly appealing to markets” [@BassolsiGardella2023]. In the case of law firms, this can be achieved by mentioning target clients (“Business” pattern) or by using words with positive meanings that attract customers (“Positive”, “Plus”, “Legal aid”, “Guarantor” patterns). The “Expert” pattern is complex, as this word can be used to indicate either a high level of service (“expert” as a synonym for “professional”) or forensic services (“expert” as a forensic scientist). It is included here into market strategy because the semantic similarity of “experts” and “forensic scientists” in the Russian language is probably used as a marketing tool by forensic companies.

@@

```{r fig-3-localmoran-maps, fig.cap="**Fig. 3.** Maps of hotspots in the spatial distribution of three law firms naming patterns: (a) Expert (b) Personal brand (c) Siberia", fig.height=10}
localmoran_maps
```

@@

The findings of the study are somewhat contradictory. On the one hand, there are 22 law firm naming patterns in Russia that can be attributed to one of the four broad SMN strategies: belonging or uniqueness, market or no-strategy. Apart from the no-strategy (~40% of firms), belonging is the most popular strategy used by 30% of law firms, followed by uniqueness (~20%) and market (~10%). On the other hand, in gerenal there are no spatial regularities in the distribution of law firms that use a particular pattern: all the parts of the country are generally the same when it comes to law firms naming. Two notable exceptions to this rule are Siberia, with a fairly popular geographical branding (“Siberia” pattern), and the European South, with an unusual spatial composition of “Expert” and “Personal brand” patterns.

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  warning = FALSE, 
  message = FALSE, 
  echo = FALSE, 
  fig.width = 6.7, 
  fig.id = FALSE,
  dpi = 1200)
IS_PAPER <- TRUE # used in script to disable saving of names and samples

# Space as thousands separator
knit_print.integer <- function(x, ...) {
  prettyNum(x, big.mark=" ")
}

registerS3method(
  "knit_print", "integer", knit_print.integer,
  envir = asNamespace("knitr")
)
```

```{r package-imports, include=FALSE}
library(dplyr)
library(knitr)
```

@@

Russia belongs to the Germanistic branch of a civil law legal system. Legal services (advice, negotiation, representation in court, etc) for organisations and individuals are usually provided by private law companies and individual lawyers. The market is divided between ordinary lawyers and advocates (members of the Bar), but there is no bar monopoly. There are ~90 000 law firms in Russia, including ~40 000 commercial law companies, ~30 000 sole practitioners, ~20 000 bar (advocate) offices, and ~10 000 bar (advocate) organisations. There are no geographical restrictions on legal practitioners, so that a law firm registered in one region of the country can provide services in any other region. There are full-service and specialised law firms, as well as Russian law firms (the majority) and foreign law firms (the influential minority).

@@

The issue of obtaining information about firm names is a challenging one. All Russian organisations are registered in the state registry operated by the Federal Tax Service (FTS of Russia). An extract from this registry about a particular company can be searched and downloaded online by any person free of charge. However, a considerable sum of money must be paid and a written request submitted to the FTS of Russia to purchase a complete copy of the registry. Commercial companies offer data resell services, but the prices are similarly high.

To address these challenges, a unified state registry of small and medium-sized enterprises (SME registry) may be used [@SMERegistry2024]. The registry has been maintained since 2016 in accordance with Article 4.1 of the Federal Law of 24 July 2007 No. 209-FZ “On the development of small and medium-sized businesses in the Russian Federation”. As a rule of thumb, all companies and individual entrepreneurs with a yearly revenue below 2 billion roubles and an average yearly number of employees below 250 are automatically included in the SME registry, without any action being required on their part. There are additional qualifications for the SME registry, such as the absence of non-SME or foreign control, but these are unlikely to affect a significant proportion of law firms. The registry is fully downloadable at the FTS open data portal at no cost. Additionally, the portal provides a monthly record of updates made since the registry's inception. To sum it up, the SME registry can be regarded as the most comprehensive free data source offering information on Russian business entities, both organisations and individuals.

@@

For a validation, a list of law firms' unique tax IDs obtained from the SME registry for the year 2023 is matched with the list of law firms' tax IDs from the Russian Financial Statements Database (RFSD) [@Bondarkov2025] for the same year. The comparison shows that ~3 300 of the firms present in the SME registry are missing from the RFSD, and ~5 200 firms present in the RFSD are missing from the SME registry. The balance of missing records is not so far from being equal. The SME registry contains ~34 000 law firms for the year 2023, so the missing rate is below 15%. Consequently, the completeness of the data is relatively high and reliable conclusions can be drawn.

@@

In this study, the term “law firm” is defined as a commercial organisation registered in Russia in accordance with the Civil Code of Russia, with a NACE Rev. 2-compatible main activity code equal to 69.10 (activity in the area of law). The aforementioned definition excludes individual entrepreneurs, foreign companies, non-profit organizations, bar (advocacy) formations, and notaries. The selection technique based on the main activity code is derived from that proposed by @Moiseeva2016. Individual entrepreneurs are excluded from the analysis because they typically operate under their personal names, and even if they have a brand, it is not listed in the SME registry. Other types of law firms are not included in the analysis primarily due to data availability issues.

@@

The Yandex Foundation Models cloud service [@Yandex2024] is used to transform the firm names into text vectors, or embeddings. The “text-search-query” model is selected because its purpose, as stated in the documentation, is to vectorize short texts. Prior to vectorization, the firm names are preprocessed to remove indications of organisational form such as “limited liability company”. This is done because such indications are common to the majority of names and thus have little value. The results of the queries to the Yandex API are stored separately for reuse, given that API calls are paid and take time to be completed. The dimension of a firm name vector is 256. Contextual embeddings (YandexGPT) were used in preference to static ones (e.g. FastText), because in the preliminary analysis, they have resulted in a much better clustering of law firm names, which is crucial for naming patterns identification.

@@

The resulting clusters are then manually grouped. A random sample of 20 names is selected from each cluster. With 50 clusters and 20 sampled names allocated in each, the total size of the dataset for manual processing is 1 000. The cluster samples are carefully examined, and a concise description is provided for each, e.g. “Names with the word “консалт/консалтинг” (consult/consulting)”. After this, the clusters are examined again and assigned to a naming pattern, e.g. “Consulting” or “Law firm”. Finally, the clusters and their naming patterns are divided into four previously mentioned naming strategies derived from Bassols i Gardella et al. [@BassolsiGardella2023]. Each cluster belongs to a single pattern, and each pattern belongs to a single strategy. Each strategy consists of several patterns, and each pattern consists of one or many clusters, forming a hierarchical structure.

@@

Spatial autocorrelation analysis is used to check whether the spatial distribution of names following each pattern identified in the previous step is random or has some regularities. Regions (constituent entities) of the Russian Federation are chosen as the unit of observation. Municipal territories, the next level of spatial division, are not used, because the exploratory data analysis has shown that the lion's share of such the spatial units do not have any law firms.

Six spatial weight matrices are utilized to ensure the robustness of the results. Three of these matrices are constructed using the K-nearest neighbors technique, where K is 3, 4 or 6. Another matrix relies on the existence of common boundaries or points to detect neighboring regions. The Sakhalin and Kaliningrad regions, which are enclaves, are manually linked to a few nearest regions. Two remaining matrices treat regions belonging to the same upper-level macro-region as neighbors, and all other regions as non-neighbors. One of these matrices uses economic regions as macro-regions, and the other uses federal districts. Economic regions are defined according to the classification that has been widely used by economic geographers and spatial development practitioners in Russia for decades, as outlined, for example, in [@Leyzerovich2010]. The composition of federal districts is defined by the Decree of the President of the Russian Federation (No. 849 of May 13, 2000) “On the Plenipotentiary Representative of the President of the Russian Federation in the Federal District”, with respective amendments.

@@

## 3.1. Law firm naming patterns and strategies

After clustering and manual grouping, 22 naming patterns are identified and assigned to one of four naming strategies (see Fig. 1 for a dendrogram and Table 1 for a sample of 10 clusters, their descriptions, patterns, and strategies). The most popular is the no-strategy followed by ~18 700 law firms. Slightly less important is the belonging strategy with ~15 000 companies. The uniqueness strategy (~7 500 law firms) is relatively rare, and the least used is the market strategy (~6 100 companies).

@@

```{r table-1-clusters}
clusters_info_table %>% 
  slice_head(n = 10) %>% 
  kable(
    col.names = c(
      "# of cluster", "# of firms", "Description",
      "Pattern", "Strategy"
    ),
    caption = "**Table 1.** Characteristics of 10 selected clusters of law firm names"
  )
```

Market strategy, according to the original definition, is used by firms “having a name directly appealing to markets” [@BassolsiGardella2023]. In the case of law firms, this can be achieved by mentioning target clients (“Business” pattern) or by using words with positive meaning that attract customers (“Positive”, “Plus”, “Legal aid”, “Guarantor” patterns). The “Expert” pattern is complex, as this word can be used to indicate either a high level of service (“expert” as a synonym for “professional”) or forensic services (“expert” as a forensic scientist). It is included here into market strategy because the semantic similarity of “experts” and “forensic scientists” in the Russian language is probably used as a marketing tool by forensic companies.

@@

```{r fig-3-localmoran-maps, fig.cap="**Fig. 3.** Maps of hotspots in the spatial distribution of three law firm naming patterns: (a) Expert (b) Personal brand (c) Siberia", fig.height=10}
localmoran_maps
```

@@

The findings of the study are somewhat contradictory. On the one hand, there are 22 law firm naming patterns in Russia that can be attributed to one of the four broad SNM strategies: belonging or uniqueness, market or no-strategy. Apart from the no-strategy (~40% of firms), belonging is the most popular strategy used by 30% of law firms, followed by uniqueness (~20%) and market (~10%). On the other hand, in gerenal there are no spatial regularities in the distribution of law firms that use a particular pattern: all the parts of the country are generally the same when it comes to law firms naming. Two notable exceptions to this rule are Siberia, with a fairly popular geographical branding (“Siberia” pattern), and the European South, with an unusual spatial composition of “Expert” and “Personal brand” patterns.
